---
title: "The LoomExperiment Classes"
author: "Daniel Van Twisk"
package: LoomExperiment
output:
  BiocStyle::html_document
vignette: >
  %\VignetteIndexEntry{An introduction to the LoomExperiment class}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r options, include=FALSE, echo=FALSE}
library(BiocStyle)
knitr::opts_chunk$set(warning=FALSE, error=FALSE, message=FALSE)
```

# The `LoomExperiment` class

## Definition

The `LoomExperiment` family of classes inherits from the main class
`LoomExperiment` as well as the Experiment class that they are named
after. For example, the `SingleCellLoomExperiment` class inherits from
both `LoomExperiment` and `SingleCellExperiment`.

The purpose of the `LoomExperiment` class is to act as an intermediary
between Bioconductor's Experiment classes and the Linnarson Lab's Loom
File Format (http://linnarssonlab.org/loompy/index.html).  The Loom
File Format uses HDF5 to store Experiment data.

The `LoomExperiment` family of classes contain the following slots.

* `colGraphs`
* `rowGraphs`

Both of these slots are `LoomGraphs` objects that describe the
`col_graph` and `row_graph` attributes as specified by the Loom File
Format.

## Create instances of LoomExperiment

There are several ways to create instances of a `LoomExperiment` class
of object.  One can plug an existing SummarizedExperiment type class
into the appropriate constructor:

```{r construct1}
library(LoomExperiment)
counts <- matrix(rpois(100, lambda = 10), ncol=10, nrow=10)
sce <- SingleCellExperiment(assays = list(counts = counts))
scle <- SingleCellLoomExperiment(sce)
## OR
scle <- LoomExperiment(sce)
```

One can also simply plug the arguments into the appropriate
constructor, since all `LoomExperiment` constructors call the
applicable class's constructor

```{r construct2}
scle <- SingleCellLoomExperiment(assays = list(counts = counts))
```

Also, it is also possible to create a `LoomExperiment` extending class
via coercion:

```{r coerce}
scle <- as(sce, "SingleCellLoomExperiment")
scle
```

Finally, one can create a `LoomExperiment` object from importing a
Loom File.

## Setting up a simple example

We will use the following `SingleCellLoomExperiment` for the remainder
of the vignette.

```{r load_l1}
l1_file <-
    system.file("extdata", "L1_DRG_20_example.loom", package = "LoomExperiment")
scle <- import(l1_file, type="SingleCellLoomExperiment")
scle
```

All the following methods apply to all `LoomExperiment` classes.

# The `LoomGraph` class

The `colGraphs` and `rowGraphs` slots of LoomExperiments correspond to
the `col_graphs` and `row_graphs` fields in the Loom File format.
Both of these slots require `LoomGraphs` objects.

A `LoomGraph` class extends the `SelfHits` class from the `S4Vectors`
package with the requirements that a `LoomGraph` object must:

* Contents must all be of class `integer` and non-negative
* Have col/row numbers corresponding to entries in the
  `LoomExperiment` object (if attached to a `LoomExperiment` object)
* May have on metadata column named 'w' that contains numeric elements

The columns `to` and `from` correspond to either `row` or `col`
indices in the `LoomExperiment` object while `w` is an optional column
that specifies the weight.

A LoomGraph can be constructed in two ways:

```{r construct_LoomGraph}
a <- c(1, 2, 3)
b <- c(3, 2, 1)
w <- c(100, 10, 1)
df <- DataFrame(a, b, w)
lg <- as(df, "LoomGraph")

## OR

lg <- LoomGraph(a, b, weight = w)
lg
```

`LoomGraph` objects can be subset by the 'row'/'col' indices.
 
```{r subset_LoomGraph}
lg[c(1, 2)]
lg[-c(2)]
```

# The `LoomGraphs` class

A `LoomGraphs` object extends the `S4Vectors:SimpleList` object.  It
contains multiple `LoomGraph` objects with its only requirement being
that it must contain `LoomGraph` objects.

It can be created simply by using `LoomGraph` objects in the
`LoomGraphs` constructor

```{r construct_LoomGraphs}
lgs <- LoomGraphs(lg, lg)
names(lgs) <- c('lg1', 'lg2')
lgs
```

# Available methods for the `LoomExperiment` 

The `LoomGraphs` assigned to these `colGraphs` and `rowGraphs` slots
can be obtained by their eponymous methods:

```{r get_col_row_graphs}
colGraphs(scle)
rowGraphs(scle)
```

The same symbols can also be used to replace the respective `LoomGraphs`

```{r replace_LoomGraphs}
colGraphs(scle) <- lgs
rowGraphs(scle) <- lgs

colGraphs(scle)
rowGraphs(scle)
colGraphs(scle)[[1]]
rowGraphs(scle)[[1]]
```

`LoomExperiment` objects can be subsetting in such a way that the
`assays`, `colGraphs`, and `rowGraphs` will all be subsetted.
`assays` will will be subsetted as any `matrix` would.  The `i`
element in the subsetting operation will subset the `rowGraphs` slot
and the `j` element in the subsetting operation will subset the
`colGraphs` slot, as we've seen from the subsetting method from
`LoomGraphs`.

```{r subset_LoomExperiment}
scle2 <- scle[c(1, 3), 1:2]
colGraphs(scle2)[[1]]
rowGraphs(scle2)[[1]]
```

```{r rbind_LoomExperiment}
scle3 <- rbind(scle, scle)
scle3
colGraphs(scle3)
rowGraphs(scle3)
colGraphs(scle3)[[1]]
rowGraphs(scle3)[[1]]
```

Finally, the `LoomExperiment` object can be exported.

```{r export_LoomExperiment}
temp <- tempfile(fileext='.loom')
export(scle2, temp)
```

# Session Info

```{r}
sessionInfo()
```
